---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning=FALSE,
  message=FALSE
)
```

# R.Scenario.Vax

<!-- badges: start -->
<!-- badges: end -->

The goal of R.Scenario.Vax is to fit a deterministic, compartmental transmission model to RSV hospitalization data and provide scenario projections for the number of hospitalizations averted due to new immunization products. 

## Installation

You can install the development version of R.Scenario.Vax from [GitHub](https://github.com/) with:


``` r
# install.packages("devtools")
remotes::install_github("chelsea-hansen/R.Scenario.Vax")
```

## Example

This is a basic example of how to prepare the population dynamics data and initial values for compartments for New York state:

```{r example}
library(R.Scenario.Vax)


ny_data = get_data(state_or_county="state",state_abbr="NY",county_name=NULL)

#Parameter Values 
print(ny_data[[1]])

```

